### 2024-04-19 paper/paper.qmd
Initial commit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(modelsummary)
library(ggplot2)
library(dplyr)
library(rstanarm)
library(parameters)
library(broom)
library(broom.mixed)
```

```{r}
#| include: false
#| warning: false
#| message: false

clean_data_below40 <- read_csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/analysis_data/analysis_data_below40.csv")|>
  select(Year.x, Cause,Population,Deaths, Death_rate_per_100000, Age)

clean_data_above40 <- read_csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/analysis_data/analysis_data_above40.csv")|>
  select(Year.x, Cause,Population,Deaths, Death_rate_per_100000, Age)
```

# Introduction

# Data {#sec-data}

## Raw Data

## Clean Data

```{r}
#| label: tbl-data-below40
#| tbl-cap: Top 10 Mortality Causes in Switzerland for young adults in 2010 
#| echo: false
#| warning: false
#| message: false

#cleaned data
head(clean_data_below40, n=10)|>
  kable(
    col.names = c("Year", "Causes", "Population", "Deaths", "Death rate", "Age"), 
    align =  c("c","c","c","c","c","c")
  )
```

```{r}
#| label: tbl-data-above40
#| tbl-cap: Top 10 Mortality Causes in Switzerland for middle ages in 2010 
#| echo: false
#| warning: false
#| message: false

#cleaned data
head(clean_data_above40, n=10)|>
  kable(
    col.names = c("Year", "Causes", "Population", "Deaths", "Death rate", "Age"), 
    align =  c("c","c","c","c","c","c")
  )
```

```{r}
#| label: tbl-common_mortality_causes_below40
#| fig-cap: Common Mortality Causes for young adults in 2010,1015,and 2019
#| echo: false

top_10_2010_below40 <- clean_data_below40$Cause[clean_data_below40$Year.x == 2010]
top_10_2015_below40 <- clean_data_below40$Cause[clean_data_below40$Year.x == 2015]
top_10_2019_below40  <- clean_data_below40$Cause[clean_data_below40$Year.x == 2019]

common_causes_below40  <- intersect(intersect(top_10_2010_below40, top_10_2015_below40), top_10_2019_below40 )
# Calculate total deaths for each common cause of death
total_deaths_below40  <- aggregate(Deaths ~ Cause, subset(clean_data_below40, Cause %in% common_causes_below40), sum)

common_causes_below40_table <- total_deaths_below40  %>% 
  arrange(desc(Deaths))%>%
  head(5)

kable(common_causes_below40_table)
```
```{r}
#| label: tbl-common_mortality_causes_above40
#| fig-cap: Common Mortality Causes for middle age adults in 2010,1015,and 2019
#| echo: false

top_10_2010_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2010]
top_10_2015_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2015]
top_10_2019_above40  <- clean_data_above40$Cause[clean_data_above40$Year.x == 2019]

common_causes_above40  <- intersect(intersect(top_10_2010_above40, top_10_2015_above40), top_10_2019_above40 )
# Calculate total deaths for each common cause of death
total_deaths_above40  <- aggregate(Deaths ~ Cause, subset(clean_data_above40, Cause %in% common_causes_above40), sum)

common_causes_above40_table <- total_deaths_above40  %>% 
  arrange(desc(Deaths))%>%
  head(5)

kable(common_causes_above40_table)
```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

datasummary(
  Deaths ~ Min + Mean + Max + SD + Var + N,
  fmt = 0,
  data = clean_data_below40
)

```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

datasummary(
  Deaths ~ Min + Mean + Max + SD + Var + N,
  fmt = 0,
  data = clean_data_above40
)
```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

clean_data_below40%>%
  filter(Cause %in% common_causes_below40_table$Cause) %>%
  ggplot(aes(x = Year.x, y = Deaths, color = Cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Number of deaths")+
  scale_x_continuous(breaks = c(2010, 2015, 2019))+ 
  ylim(0,250)# Adjust the limits as needed
```

```{r}
#| label: fig-abc
#| fig-cap: abc
#| echo: false
#| warning: false
#| message: false

clean_data_above40%>%
  filter(Cause %in% common_causes_above40_table$Cause) %>%
  ggplot(aes(x = Year.x, y = Deaths, color = Cause)) +
  geom_line() +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Year", y = "Number of deaths")+
  scale_x_continuous(breaks = c(2010, 2015, 2019))+ 
  ylim(100,650)# Adjust the limits as needed
```

# Model

## Model set-up

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

### Model justification

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

## Load models
cause_of_death_below40 <-
  readRDS(file ="~/Top-mortality-causes-of-age-groups-in-Switzerland/models/cause_of_death_below40.rds")
cause_of_death_above40 <-
  readRDS(file ="~/Top-mortality-causes-of-age-groups-in-Switzerland/models/cause_of_death_above40.rds")

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary(
  list(
    " Below 40" = cause_of_death_below40
))
```
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary(
  list(
    " Above 40" = cause_of_death_above40
))
```
```

# Discussion

## First discussion point {#sec-first-point}

```{r}

```

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-19 scripts/02-data_cleaning.R

@@ -1,0 +1,246 @@
+#### Preamble ####
+# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Author: 
+# Date:
+# Contact: thu.dong@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: Install needed packages
+# Any other information needed? 
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+
+#### Clean data ####
+
+# read data in the raw data files (below age 40)
+switzerland_2010_20to24 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_20to24.csv" )
+switzerland_2010_25to29 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_25to29.csv" )
+switzerland_2010_30to34 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_30to34.csv" )
+switzerland_2010_35to39 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_35to39.csv" )
+
+switzerland_2015_20to24 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_20to24.csv" )
+switzerland_2015_25to29 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_25to29.csv" )
+switzerland_2015_30to34 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_30to34.csv" )
+switzerland_2015_35to39 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_35to39.csv" )
+
+switzerland_2019_20to24 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2019_20to24.csv" )
+switzerland_2019_25to29 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2019_25to29.csv" )
[…]

### 2024-04-19 scripts/04-model.R

@@ -1,0 +1,86 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(parameters)
+library(broom)
+library(broom.mixed)
+
+#### Read data ####
+analysis_data_below40 <- read_csv("~/Top-mortality-causes-of-age-groups-in-Switzerland/data/analysis_data/analysis_data_below40.csv")
+analysis_data_above40 <- read_csv("~/Top-mortality-causes-of-age-groups-in-Switzerland/data/analysis_data/analysis_data_above40.csv")
+### Model data ####
+
+# get common causes of deaths for both age group
+
+top_10_2010_below40 <- analysis_data_below40$Cause[analysis_data_below40$Year.x == 2010]
+top_10_2015_below40 <- analysis_data_below40$Cause[analysis_data_below40$Year.x == 2015]
+top_10_2019_below40 <- analysis_data_below40$Cause[analysis_data_below40$Year.x == 2019]
+
+common_causes_below40 <- intersect(intersect(top_10_2010_below40, top_10_2015_below40), top_10_2019_below40) 
+total_deaths_below40 <- aggregate(Deaths ~ Cause, subset(clean_data_below40, Cause %in% common_causes_below40), sum)
+
+common_causes_below40_table <- total_deaths_below40 %>% 
+  arrange(desc(Deaths))%>%
+  head(5)
+
+data_below40 <-
+  analysis_data_below40 |>
+  filter(Cause %in% common_causes_below40_table$Cause) 
+
+top_10_2010_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2010]
+top_10_2015_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2015]
+top_10_2019_above40  <- clean_data_above40$Cause[clean_data_above40$Year.x == 2019]
+
+common_causes_above40  <- intersect(intersect(top_10_2010_above40, top_10_2015_above40), top_10_2019_above40 )
[…]

### 2024-04-19 paper/paper.qmd
Commit introduction and abstract

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Top Mortality causes in young adults and middle age adults in Switzerland"
+subtitle: "Analyzing data from the WHO database"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Thu Dong
+thanks: "Code and data are available at: https://github.com/thu2912/Top-mortality-causes-of-age-groups-in-Switzerland.git"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study employs Bayesian analysis to investigate mortality causes among young adults (aged 20 to below 40) and middle-aged adults (aged 40 to below 60) in Switzerland.
+Analyzing mortality data from national registries and healthcare databases, significant disparities in mortality causes between the two age groups are identified.
+The findings underscore the importance of targeted interventions and policy measures to address the distinct health needs of young and middle-aged adults.
+This research contributes to advancing understanding of mortality dynamics in Switzerland and informs evidence-based public health strategies."
 format: pdf
@@
 # Introduction
+In recent years, understanding the mortality causes among different age groups has become increasingly crucial for public health research and policy-making.
+Mortality patterns can vary significantly across age cohorts, reflecting not only the natural aging process but also the diverse lifestyle factors, socioeconomic conditions, and healthcare accessibility experienced by different demographic groups.
+In Switzerland, a country renowned for its high-quality healthcare system and favorable living standards, investigating mortality causes among young and middle-aged adults presents a compelling avenue for exploring the health dynamics within these cohorts.
+
+This research aims to delve into the mortality causes prevalent among young adults (aged 20 to below 40) and middle-aged adults (aged 40 to below 60) in Switzerland, employing a Bayesian analytical approach.
+While numerous studies have investigated mortality trends in broader age categories or specific disease cohorts, there remains a paucity of research focusing explicitly on these two distinct age groups.
+By utilizing Bayesian methods, this study offers a novel perspective on mortality analysis, allowing for the incorporation of prior knowledge, uncertainty quantification, and robust inference in modeling mortality causes.
+
+The findings of this research shed light on the stark differences in the top mortality causes between young and middle-aged adults in Switzerland.
+By examining comprehensive mortality data from national registries and healthcare databases, this study identifies and analyzes the leading causes of death within each age cohort, elucidating the underlying factors contributing to these disparities.
+Moreover, the Bayesian framework enables the exploration of uncertainty and variability in mortality patterns, providing valuable insights for policymakers, healthcare practitioners, and researchers in developing targeted interventions and preventive strategies tailored to the specific needs of each age group.
+
+Through this research, we seek to contribute to the broader understanding of mortality dynamics in Switzerland and advance the discourse on public health initiatives aimed at reducing premature deaths and improving overall population health.
+By uncovering the distinct mortality landscapes among young and middle-aged adults, we aim to inform evidence-based policymaking and resource allocation strategies that address the unique health challenges faced by these demographic cohorts.
 
@@
 ```{r}
-#| label: fig-abc
-#| fig-cap: abc
+# #| label: fig-abc
+# #| fig-cap: abc
 #| echo: false
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+# #| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-modelresults1
+# #| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-04-20 paper/datasheet.qmd
Update datasheets

---
title: "Datasheet for 'analysis_data_above40'and 'analysis_data_below40'"
author: 
  - Thu Dong
thanks: "Code and data are available at: https://github.com/thu2912/Top-mortality-causes-of-age-groups-in-Switzerland.git."
date: today
date-format: long
abstract: "Datasheet for 'analysis_data_above40' and 'analysis_data_below40', looking into the motivation, collection process, preprocessing/cleaning/labelling, distribution, and maintenance of it"
format: pdf
number-sections: true
bibliography: references.bib
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable an analysis of the top mortality causes in Switzerland in young adults and middle-aged adults using the data that is provided in the WHO Mortality database. We were unable to find a way to a data set that has already separated the age groups, young adults and middle-aged adults.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset for analysis was created by Thu Dong
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
No funds were needed nor granted for the dataset that was used for the analysis.
4. *Any other comments?*
    - NA

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- A row in the dataset refers to a cause of death in a specific year for a specific age group 
2. *How many instances are there in total (of each type, if appropriate)?*
	- There were 30 instances in each dataset and there were 2 different datasets. 
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- It is a smaller sample from the big WHO Mortality database. The larger dataset all causes of death in 120 countries. The data sets contain only the top 10 number of deaths per age group for the years 2010, 2015, and 2019.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- It consists of an integer value of the annual number of deaths of each corresponding cause, population, death rate, and age. 
5. *Is there a label or target associated with each instance? If so, please provide a description.*
yes. It is unique causes in unique years 
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information but might include, for example, redacted text.*
	- No there is no information missing. 
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, and social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, unique causes in the year
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, and other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The data is self-contained.  
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No
16. *Any other comments?*
	- NA

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was obtained by cleaning the raw data. The raw data was obtained from the WHO Mortality Database. The data was not reported by subjects or indirectly inferred/derived from other data. 
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data was collected by downloading manually through the WHO Mortality database
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is a sample from a larger set as we use certain age groups given from the WHO database.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Thu Dong was the only person involved in the data collection process. 
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected during April 2024. The raw data was only available from 2000 to 2019.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please describe these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No ethical review process was conducted
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was directly downloaded from the WHO official website. 
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- There was no direct involvement of individuals in obtaining the raw data.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Data is available publicly. Therefore, there is no need for any approval.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- NA. 
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- NA. 
12. *Any other comments?*
	- NA

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes. Only selected needed columns from the website and merged data from different age groups and years. 
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- NA
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- NA
4. *Any other comments?*
	- NA

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- It has been used to analyze top mortality causes in young adults and middle-aged adults in Switzerland by creating the Negative Binomial Model.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- https://github.com/thu2912/Top-mortality-causes-of-age-groups-in-Switzerland.git
3. *What (other) tasks could the dataset be used for?*
	- One can use this dataset for other models such as simple regression. 
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No.
6. *Any other comments?*
	- NA

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No 
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset will be distributed on GitHub. It doesn't have a DOI.
3. *When will the dataset be distributed?*
	- April 20, 2024
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No
6. *Do any export controls or other regulatory restrictions apply to the dataset or individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No
7. *Any other comments?*
	- NA

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Thu Dong 
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	-thu.dong@mail.utoronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Very likely it will not happen. In the case it does, it may happen once a year, and it will be communicated through GitHub
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	-No the dataset is just updated. Although a history is available through GitHub.

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Pull request from Github. 
8. *Any other comments?*
	- NA

\newpage

# References

### 2024-04-20 scripts/00-simulate_data.R
Update Scripts

@@ -1,19 +1,28 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates data
+# Author: Thu Dong
+# Date: 20 April 2024
+# Contact: thu.dong@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Install needed packages
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(tibble)
+library(dplyr)
 
-#### Simulate data ####
-# [...ADD CODE HERE...]
+# Define the causes, years, age groups, and base population
+simulation_data <- tibble(
+  cause = rep(x = c("Heart", "Stroke", "Cancer", "Tuberculosis"), each = 3),
+  year = rep(x = c(2010, 2015, 2019), times = 4),
+  deaths = rnbinom(n = 12, size = 40, prob = 0.5),
+  age = rep(c("20-39", "40- 59"), length.out = 12),
+  population = rep(c(500000, 600000, 700000), length.out = 12)
+)
+
+
+
 
 
 

### 2024-04-20 scripts/01-download_data.R

@@ -1,26 +1,43 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from WHO Mortality Database
+# Author: Thu Dong
+# Date: 20 April 2024
+# Contact: thu.dong@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+### The data is downloaded directly from WHO Mortality Database 
+# read data in the raw data files (below age 40)
+switzerland_2010_20to24 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_20to24.csv" )
+switzerland_2010_25to29 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_25to29.csv" )
+switzerland_2010_30to34 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_30to34.csv" )
+switzerland_2010_35to39 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2010_35to39.csv" )
 
+switzerland_2015_20to24 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_20to24.csv" )
+switzerland_2015_25to29 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_25to29.csv" )
+switzerland_2015_30to34 <- read.csv( "~/Top-mortality-causes-of-age-groups-in-Switzerland/data/raw_data/switzerland_2015_30to34.csv" )
[…]

### 2024-04-20 scripts/02-data_cleaning.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: 
-# Date:
+# Purpose: Cleans the raw data from the WHO Mortality website
+# Author: Thu Dong
+# Date: 20 APril 2024
 # Contact: thu.dong@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: Install needed packages
-# Any other information needed? 
+
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-04-20 scripts/03-test_data.R

@@ -1,15 +1,42 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests to check if the cleaned data returns TRUE for given conditions
+# Author: Thu Dong
+# Date: today
+# Contact: thu.dong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+
+# Tests if death column is integers
+analysis_data_above40$Deaths |> class() == "integer"
+analysis_data_below40$Deaths |> class() == "integer"
+
+# Tests if there are four different years
+length(unique(analysis_data_below40$Year)) == 3
+length(unique(analysis_data_above40$Year)) == 3
+
+# Tests if number of deaths is larger than 0
+all(analysis_data_below40$Deaths > 0)
+all(analysis_data_above40$Deaths > 0)
+
+# Tests if death rate is larger than 0
+all(analysis_data_above40$Death_rate_per_100000 > 0)
+all(analysis_data_below40$Death_rate_per_100000 > 0)
+
+# Tests if death rate is in integers
+analysis_data_below40$Death_rate_per_100000 |> class() == "numeric"
+analysis_data_above$Death_rate_per_100000 |> class() == "numeric"
+
+# test if each year we collected top 10 causes 
+year_data <- analysis_data_above40[analysis_data_above40$Year.x == yr, ]
+
+year_data_2 <- analysis_data_below40[analysis_data_below40$Year.x == yr, ]
+
+length(unique(year_data$Causes)) == 10 
+
+length(unique(year_data_2$Causes)) == 10 

### 2024-04-21 paper/paper.qmd
Update datasheets
Update Introduction, model, results, discussion

--- 2024-04-19 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Top Mortality causes in young adults and middle age adults in Switzerland"
+subtitle: "Analyzing Switzerland Mortality causes data in 2010, 2015, adn 2019 from WHO database"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Thu Dong
+thanks: "Code and data are available at: https://github.com/thu2912/Top-mortality-causes-of-age-groups-in-Switzerland.git"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Motivated by the lack of findings on different mortality causes for different age groups, this paper examines the main causes of mortality among young adults and middle-aged adults in Switzerland using data from the WHO Mortality Database for the years 2010, 2015, and 2019.
+The paper found that the leading causes of death for young adults are self-harm and drug use disorders, while middle-aged adults are most affected by ischemic heart disease and trachea, bronchus, and lung cancers.
+Moreover, it is also found that one of the main mortality causes in both age groups is self-harm which highlights the importance of mental health."
 format: pdf
@@
 # Introduction
+Mortality study has become increasingly important in modern society.
+Understanding mortality causes is vital for medicine and health research, as well as implementing policies that can help increase the quality of living and life expectancy.
+With Switzerland being the country with the 4th highest life expectancy in the world, there has been much research about the causes of death in Switzerland.
+However, there have not been a lot of studies that focus on the different causes of death in young adults (20 to 39) and middle-aged adults (40 to 59) in the country.
+
+Therefore, in this paper, we will investigate the top mortality causes in young adults and middle-aged adults in Switzerland and discuss some trends and patterns in healthcare that were witnessed in the data.
+The estimand that we’re trying to estimate is the number of deaths by different mortality causes in young adults and middle-aged adults.
+We aim to provide empirical evidence on different causes of death in two age groups and investigate the reasons behind the differences in health issues in two age groups.
+This paper uses data obtained from the open database, WHO Mortality database, which exhibits information on mortality causes of deaths filtered by country, sex, year, and age group.
+The data used in this paper were the Mortality causes in Switzerland in the years 2010,2015, and 2019 as it gives a general idea of the mortality causes from 2010 to 2019, filtered by age group.
+
+We found that on average, the top causes of death in young adults are self-harm and drug use disorder, and the top mortality causes in middle-aged adults are Ischaemic heart disease, as well as trachea, bronchus, and lung cancers.
+Additionally, we found a trend of high rate of self-harm in both young adults and middle-aged adults.
+The findings of the paper highlight the importance of healthy habits such as a healthy diet and exercise and the reduction of harmful substance use, such as tobacco, to maintain a healthy heart and lungs.
+Additionally, the paper also highlights the importance of mental health and policies that allow for accessible mental health support.
+Moreover, this paper’s findings demonstrate how policymakers play a major role in addressing public health issues through effective interventions.
+
+The remainder of the paper is structured as follows: @sec-data introduces the data used for analysis and findings, including visualizations of the variables of interest, @sec-model presents the Negative Binomial regression model to justify the relationship between the number of deaths and causes of deaths, @sec-result displays the interpretations of the model alongside other findings from analyzing the data, and @sec-discussion provides a discussion on the implications of the findings as well as the weaknesses of this paper and its next steps for further study on this subject.
 
@@
 
-## Raw Data
+## Data Sources
+The data utilized in this paper is provided by the  World Health Organization  WHO) specifically from the WHO Mortality Database (“Global Health Estimates: Leading Causes of Death” 2021) which focuses on health estimates and global mortality.
+At present, the WHO Mortality database contains data concerning top mortality causes in 120 different countries from 2000 until 2019, covering large age groups from 0 to above 85.
+WHO is a data source that offers openly available data to the public and can be downloaded free from the website.
+Key elements of the WHO Mortality Database include country-specific data categorized by age group, sex, and year; detailed data on causes of death by number of deaths; a visualization portal for data exploration which enhanced user experience; and raw data files for download.
+
+The database’s data collection methods include collaborating with health authorities and statistical agencies from over 120 countries.
+One of the steps in the data collection process is using the Civil Registration and Vital Statistics (CRVS) Systems as it records events like deaths, births, causes of deaths, etc which is reported and maintained by national health authorities.
+Another step of collecting data is processing data with death data obtained by count and population size by sex and age estimated by using census data and annual population estimates.
+Other steps after collecting the data are standardized coding and quality assurance.
+Although different steps of the collection have been highlighted, the specifics of each step were not available publicly.
 
 ## Clean Data
+he data utilized for this paper is downloaded manually through the WHO Mortality Database.
+All the data analysis was done through R programming language (@citeR) with the aid of the following packages:rstanarm (@rstanarm), tidyverse (@tidyverse), knitr (@knitr), modelsummary (@modelsummary), ggplot2 (@ggplot2), dplyr (@dplyr), parameters (@Parameters), broom (@broom), broom.mixed (@broommixed).
+
+The data is prepared specifically by downloading data for each age group separately (20 to under 40 and 40 to under 60).
+Then, the variables selected for the analysis are year, population, number of deaths, and causes of death.
+Additionally, two other variables were created: age which takes values ‘below 40’ and ‘above 40’ to respectively refer to young adults and middle-aged adults.
+Overall, 2 different data sets were created based on age groups.
+For each age group, data was then merged from the years 2010, 2015, and 2019.
+This resulted in two data sets, one for young adults, and one for middle-aged adults.
+Finally, I ensured consistency in variable naming and format during the merging process.
+This two data sets are created to analyze trends in mortality causes among young and middle-aged adults in Switzerland using data from 2010, 2015, and 2019.
+The two analysis datasets include 30 observations each and 6 variables each: Year, deaths, causes, death rate, and age.
+Each observation shows one of the mortality causes in one of the 3 years researched.
+For example, @tbl-table1 and @tbl-table2 display the top 10 causes of death in Switzerland in 2010.
+In both tables, year recorded the years that this data was reported from, causes recorded some mortality causes such as self-harm, drug use disorders, road injury, breast cancer, etc, the population is constant for an age group in a year, deaths recorded the number of deaths attributed to each cause within the population.
+Furthermore, the death rate is the number of deaths per 100,000 population for each cause, calculated based on the number of deaths over population, and the age variable reports the age category which is "below 40" when looking at young adults (20 to under 40), and “above 40" for middle age adults (40 to under 60).
+Similar datasets with the same attributes can be used for this analysis.
+However, since we were unable to find data that have separated observations into each age group.
+Therefore, the data that is used for this paper is created using data from WHO Mortality Database.
 
 ```{r}
-#| label: tbl-data-below40
+#| label: tbl-table1
 #| tbl-cap: Top 10 Mortality Causes in Switzerland for young adults in 2010 
@@
 #| message: false
-
+#| tbl-pos: H
 #cleaned data
@@
 ```{r}
-#| label: tbl-data-above40
+#| label: tbl-table2
 #| tbl-cap: Top 10 Mortality Causes in Switzerland for middle ages in 2010 
@@
 
+## Variable of interest. 
+### Causes of death among young adults in Switzerland.
+
+@tbl-table3 shows the common causes of death in 2010, 2015, and 2019 and the total number of deaths per cause in young adults in Switzerland.
+It is shown that in those 3 years, there were 595 deaths caused by Self-harm, 260 deaths caused by Drug use disorders, 181 deaths caused by Road injury, 79 deaths caused by Brain and nervous system cancers, and 73 deaths caused by Falls.
+
+@fig-fig1 shows the line plot that illustrates the trend of the number of deaths for common causes among young adults over the years 2010, 2015, and 2019.
+Self-harm appears to have the highest number of deaths across all three years, with a slight decrease from 2010 to 2019.
+Drug use disorders also show a fairly constant number of deaths from 2010 to 2015, followed by a slight decline by 2019.
+Additionally, Road injuries show a stable decrease in the number of deaths throughout the 3 years, from approximately 75 in 2010 to around 40 in 2019.
+Brain and nervous system cancers and Falls show a constant low number of deaths in both, at approximately 25 cases in each in all 3 years.
+
 ```{r}
-#| label: tbl-common_mortality_causes_below40
-#| fig-cap: Common Mortality Causes for young adults in 2010,1015,and 2019
+#| label: tbl-table3
+#| tbl-cap: Common Mortality Causes for young adults in 2010,1015,and 2019
 #| echo: false
-
+#| tbl-pos: H
 top_10_2010_below40 <- clean_data_below40$Cause[clean_data_below40$Year.x == 2010]
@@
 ```
+
 ```{r}
-#| label: tbl-common_mortality_causes_above40
-#| fig-cap: Common Mortality Causes for middle age adults in 2010,1015,and 2019
+#| label: fig-fig1
+#| fig-cap:  Common mortality causes in young adults in 2010, 2015, and 2019
 #| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: H
 
+clean_data_below40%>%
+  filter(Cause %in% common_causes_below40_table$Cause) %>%
+  ggplot(aes(x = Year.x, y = Deaths, color = Cause)) +
+  geom_line() +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  labs(x = "Year", y = "Number of deaths")+
+  scale_x_continuous(breaks = c(2010, 2015, 2019))+ 
+  ylim(0,250)# Adjust the limits as needed
+```
+
+### Causes of death among Middle-aged adults in Switzerland. 
+@tbl-tbl4 shows the common causes of death in 2010, 2015, and 2019 and the total number of deaths per cause in middle-aged adults in Switzerland.
+It is shown that in those 3 years, there were 1656 deaths caused by Ischaemic heart disease, 1548 deaths caused by Trachea, bronchus, lung cancers, 1285 deaths caused by Self-harm, 825 deaths caused by Breast cancer, and 618 deaths caused by Cirrhosis of the liver.
+
+```{r}
+#| label: tbl-tbl4
+#| tbl-cap: Common Mortality Causes for middle age adults in 2010,1015,and 2019
+#| echo: false
+#| tbl-pos: H
 top_10_2010_above40 <- clean_data_above40$Cause[clean_data_above40$Year.x == 2010]
@@
 
+@fig-fig2 is the line plot that illustrates the number of deaths for common causes among middle-aged adults in 2010, 2015, and 2019.
+Ischaemic heart disease appears to have the highest number of deaths across all three years.
+However, there was a significant decrease in the number of deaths caused by Ischaemic heart disease from 2010 to 2015, followed by a slight decrease from 2015 to 2019.
+Compared to 2019, there were over 100 more deaths caused by Ischaemic heart disease in 2010.
+Trachea, bronchus, and lung cancers constantly decreased in the number of deaths throughout the 3 years, from approximately 580 in 2010 to around 480 in 2019.
+Self-harm also shows a fairly constant number of deaths in those 3 years, at approximately over 400.
+Additionally, Breast cancer demonstrates a slight decrease in the number of deaths from 2010 to 2019.
+There was also a small decrease in the number of deaths from Cirrhosis of the liver from 2010 to 2015 and remained constant in 2019.
+
 ```{r}
-#| label: fig-abc
-#| fig-cap: abc
+#| label: fig-fig2
+#| fig-cap: Common mortality causes in middle-aged adults in 2010, 2015, and 2019
 #| echo: false
@@
 #| message: false
-
-datasummary(
-  Deaths ~ Min + Mean + Max + SD + Var + N,
-  fmt = 0,
-  data = clean_data_below40
-)
-
-```
-
-```{r}
-#| label: fig-abc
-#| fig-cap: abc
-#| echo: false
-#| warning: false
-#| message: false
-
-datasummary(
-  Deaths ~ Min + Mean + Max + SD + Var + N,
-  fmt = 0,
-  data = clean_data_above40
-)
-```
-
-```{r}
-#| label: fig-abc
-#| fig-cap: abc
-#| echo: false
-#| warning: false
-#| message: false
-
-clean_data_below40%>%
-  filter(Cause %in% common_causes_below40_table$Cause) %>%
-  ggplot(aes(x = Year.x, y = Deaths, color = Cause)) +
-  geom_line() +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  labs(x = "Year", y = "Number of deaths")+
-  scale_x_continuous(breaks = c(2010, 2015, 2019))+ 
-  ylim(0,250)# Adjust the limits as needed
-```
-
-```{r}
-#| label: fig-abc
-#| fig-cap: abc
-#| echo: false
-#| warning: false
-#| message: false
-
+#| fig-pos: H
 clean_data_above40%>%
@@
 
-# Model
+### Basic Summary statistics
+
+@tbl-sum1 is the summary statistics of @tbl-table1 showing its minimum, mean, maximum, standard deviation, variance, and sample size of young adults’ top mortality causes data in 2010, 2015, and 2019.
+The summary statistics show that the mean number of deaths is 50, with a minimum of 14, and a maximum of 204 for a sample size of 30.
+The variance is high because the range of the data is large as the variance is 3077.
+
+```{r}
+#| label: tbl-sum1
+#| tbl-cap: Summary statistics for young adults
+#| echo: false
+#| warning: false
+#| message: false
+#| tbl-pos: H
+datasummary(
+  Deaths ~ Min + Mean + Max + SD + Var + N,
+  fmt = 0,
+  data = clean_data_below40
+)
+
+```
+
+```{r}
+#| label: tbl-sum2
+#| tbl-cap: Summary statistics for middle-aged adults
+#| echo: false
+#| warning: false
+#| message: false
+#| tbl-pos: H
+
+datasummary(
+  Deaths ~ Min + Mean + Max + SD + Var + N,
+  fmt = 0,
+  data = clean_data_above40
+)
+```
+
+@tbl-sum2 is the summary statistics of @tbl-table2 showing its minimum, mean, maximum, standard deviation, variance, and sample size of middle-aged adults’ top mortality causes data in 2010, 2015, and 2019.
+The summary statistics show that the mean number of deaths is 272, with a minimum of 103, and a maximum of 603 for a sample size of 30.
+The variance is high because the range of the data is large as the variance is 26240.
+
+The variable in the data set is measured by using the results of the data collection in the WHO Mortality Database.
+The number of deaths is measured in integers and by the total number of deaths reported to the national database that year.
+Population is also an integer that is measured using the demographic data of the country.
+Additionally, Causes is a categorical variable that records all popular causes that lead to death in a year.
+Age group is a categorical variable that records whether or not a person is “above 40” or “below 40”.
+While “below 40” measures young adults, “above 40” measures middle-aged adults.
+Finally, death rate per 100,000 is measured by dividing the total number of death of a specific mortality causes by the population and times 100,000.
+
+# Model {#sec-model}
 
 ## Model set-up
+Understanding the impact of different causes of death on mortality rates can allow policymakers to create policies that target certain alarming health issues and reduce the death rates in the country.
+Furthermore,  policymakers and researchers can prioritize efforts to address the leading causes of mortality and allocate resources effectively.
+Therefore, to further understand the relationship between causes of death and death rates, we construct a negative binomial regression with the Bayesian framework:
 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i &\sim \text{NegBin}(\mu_i, \sigma) \\
+\log(\mu_i) &= \beta_0 + \beta_1 x_i \\
+\beta_0 &\sim \text{Normal}(0, 2.5) \\
+\beta_1 &\sim \text{Normal}(0, 2.5) \\
+\sigma &\sim \text{Exponential}(1)
 \end{align}
+
+- $y_i$ is the number of deaths in young adults or middle aged adultsin Switzerland
+- $\beta_0$ is the coefficient of intercept 
+- $\beta_1$ is the coefficient for an average increase in the number of death of a mortality cause
+- $x_i$ is the causes of death in young adults or middle aged adults
 
@@
 
-# Results
+A negative binomial is used to analyze the relationship between the number of deaths and the causes of death in each age group.
+We assume that there is a relationship between the causes of death and the number of deaths which was somewhat demonstrated in Figure 1 and Figure 2.
+The reason why we use a negative binomial regression to analyze this relationship is that negative binomial regression is suitable for counting data such as the number of deaths.
+Moreover, it takes into account overdispersion which happens when variance exceeds the mean which is shown in the summary statistics.
+Therefore, we pick the Negative Binomial regression over the Poisson regression.
+Additionally, we apply the Bayesian framework to our analysis as it allows for prior knowledge to be included in the model.
+Therefore, we assume that all coefficient follows a normal distribution with a mean of 0 and a standard deviation of 2.5 which means we allow some variation in the coefficient of the model.
 
-Our results are summarized in @tbl-modelresults.
+# Results {#sec-result}
+## Model 1
+
+@tbl-modelresults1 shows the result of the relationship between the cause of death and the number of deaths for young adults in Switzerland.
+The intercept represents the expected log count when a young adult has Brain and nervous system cancers which is approximately 3.283.
+The expected log count of deaths among young adults with drug use disorder increases by 1.192 which is associated with more death for individuals with drug disorder compared to individuals with Brain and nervous system cancers.
+Additionally, young individuals who passed away because of a fall are associated with a 0.090 decrease in the expected log count compared to young individuals with Brain and nervous system cancers which means that deaths because of Falling in youth are associated with fewer deaths than Brain ner.
+Notably, it is witnessed that more young adults pass away due to Road injuries than Brain and nervous system cancer with a 0.832 increase in the expected log count.
+Finally, young adults who have self-harm tendencies have a 2.019 increase in the expected log count compared to Brain and nervous system cancer which indicates that self-harm is associated with a higher number of deaths in young adults.
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-modelresults1
+#| tbl-cap: "Relationship between causes of deaths and number of deaths in young adults" 
 #| warning: false
-
+#| tbl-pos: H
 modelsummary(
@@
 ```
+
+## Model 2
+
+@tbl-modelresults2 shows the result of the relationship between the cause of death and the number of deaths for middle-aged adults in Switzerland.
+The intercept represents the expected log count when a young adult has Breast cancer which is approximately 5.634.
+This indicates that on average, breast cancer is associated with 279.77 deaths in a year.
+The expected log count of middle-aged adults with Self-harm increases by 0.445 which indicates that self-harm is associated with more death than Breast cancer in middle-aged adults.
+Additionally, middle-aged individuals who passed away because of Cirrhosis of the liver are associated with a 0.277 decrease in the expected count compared to middle-aged individuals with Breast Cancer.
+Notably, it appears that more middle-aged adults pass away due to Trachea, bronchus, and lung cancers than Breast cancer with a 0.629 increase in the expected log count.
+Finally, middle-aged adults who have Ischaemic heart disease have a 0.694 increase in the expected log count compared to Breast cancer.
+This indicates that Trachea, bronchus, lung cancer, and Ischaemic heart disease are associated with a higher number of deaths in middle-aged adults.
+
 ```{r}
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-modelresults2
+#| tbl-cap: "Relationship between causes of deaths and number of deaths in middle-aged adults" 
 #| warning: false
-
+#| tbl-pos: H
 modelsummary(
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+## The differences between mortality causes of young adults and middle-aged adult
 
-```{r}
+After analyzing the data, it appears that the number of deaths in middle-aged adults is much higher than in young adults.
+For example, @tbl-tbl3 and table—tbl4 shows that in 2010,2015, and 2019, the number of deaths by Ischaemic heart disease and Trachea, bronchus, and lung cancers in middle-aged adults is over 3 times the number of deaths by Self-harm in young adults.
+witnessed that top mortality causes are vastly different between young adults and middle-aged adults.
 
-```
+@fig-fig2 demonstrates that most of the top causes of death in middle-aged adults are noncommunicable diseases such as Ischaemic heart disease, Trachea, bronchus, and lung cancers,  Breast cancer, and Cirrhosis of the liver.
+This number indicates that the number of deaths that are caused by a decline in heart and lung health skyrockets as soon as they become middle-aged adults as diseases like Ischaemic heart disease, Trachea, bronchus, and lung cancers were not among the top mortality causes in young adults.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+While middle-aged adults experience physical health decline, it appears that deaths caused by mental health issues such as Self-harm and Drug use disorder, and injuries such as road injuries are the main causes of deaths in young adults.
+Notably, the two main causes of death in young adults are self-harm and drug use disorder with a total of approximately 600 cases of Self-harm and 260 cases of drug use disorder.
+One of the reasons why young adults tend to experience fewer noncommunicable diseases is the resilience in their youthful bodies and healthy habits of diet and regular exercise.
 
-## Second discussion point
+## Leading cause of death in middle-aged people in Switzerland. 
+As shown in @tbl-tbl4, Ischaemic heart disease (IHD) is the leading cause of death in middle-aged adults.
+Notably, the leading cause of death for the whole population in Switzerland is Ischaemic heart disease with approximately 11,000 cases per year since 2010.
+(“Coronary Heart Disease in Switzerland” 2020).
+This disease is not only common in Switzerland, but it is also common worldwide, It results from reduced blood flow to the heart muscle due to narrowed or blocked coronary arteries.
+Some risk factors for this disease include high blood pressure, high cholesterol, smoking, diabetes, and obesity.
+It is reported that IHD is responsible for approximately 16% of total deaths worldwide1.(World 2020).
 
-## Third discussion point
+One of the reasons why IHD is the highest cause of death in Switzerland is because of the citizen’s tobacco usage.
+It is known that tobacco usage is one of the risk factors for IHD.
+In 2017, Swiss smokers used an average of 9.5 cigarettes per day.
+Additionally, approximately 21% of smokers consumed at least 20 cigarettes per day.
+Overall, 6% of the Swiss population aged 15 and over were smokers.
+(“Level of Use - Swiss Association for Tobacco Control” 2016).
+It is shown that smokers are 2 or 4 times more likely to develop IHD than non-smokers.
+With these statistics in mind, policymakers passed regulations such as banning smoking indoors and banning selling tobacco to people under 18.
+
+## Self-harm is one of the prevalent causes of death in Switzerland
+ Among the top mortality causes in young adults and middle-aged adults, Self-harm is the common cause of death in both age groups.
+As shown in @fig-fig1 and @fig-fig2 , self-harm is the leading cause of death among young adults in Switzerland and it is the 3rd most common mortality cause in middle-aged adults.
+Notably, the number of deaths caused by self-harm in middle-aged adults is higher than in young adults.
+@fig-fig1 and @fig-fig2 demonstrate that on average, there were approximately 200 cases of deaths by self-harm in young adults in a year while there were approximately 400 deaths by self-harm in middle-aged adults in a year on average.
+This illustrates that self-harm is a concerning issue in Switzerland and understanding its underlying factors is crucial to prevent it.
+
+Switzerland witnessed an alarming rate of self-harm incidents.
+A government-commissioned study in the French-speaking part of Switzerland documented 554 cases of self-harm by 490 patients over a ten-month period.
+Additionally, half of the patients were aged 18 to 34, and 65% of them were reported to have a difficult socioeconomic situation.
+(SWI swissinfo.ch 2019).
+In 2016, Switzerland launched a national action plan on self- harm/ suicide prevention with some key measures: raise awareness about these issues, provide accessible help, recognize early behavior of self-harm and timely intervene, reduce means of suicide and self-harm, etc.
 
@@
 
-Weaknesses and next steps should also be included.
+### Weaknesses
 
-\newpage
+A limitation that affects external validation of the analysis is that since we use data from Switzerland in specific age groups, we can make the same conclusion for the same age group in another country.
+Moreover, the number of years used in the research is limited.
+Therefore, the results could be inconclusive.
+Additionally, another limitation is that data from the WHO Mortality Database is only available from 2000 to 2019, we are unsure of the mortality causes in recent years.
+Moreover, a limitation regarding the research's internal validity is that by using the Negative Binomial models, we assume a previously existing relationship between the number of deaths and causes of death.
+Additionally, since we’re using the Negative Binomial models, we may missed other models that may provide better results that we have not considered.
 
-\appendix
+### Next step
 
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
+In the future, we will extend the study to include more years and possibly conduct a time-series analysis to observe trends over a continuous period.
+This change can also help us study for effect of different policies on mortality rates.
+Additionally, we can also compare the mortality causes and trends in Switzerland with those in other countries with similar socioeconomic conditions.
+As a result, we can identify the unique factors that make Switzerland one of the countries with the longest life expectancy.
 

### 2024-04-21 scripts/04-model.R
Update model script

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models for young adults and middle-aged adults
+# Author: Thu Dong
+# Date: 20 April, 2024
+# Contact: thu.dong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+
 
 
 #### Workspace setup ####

### 2024-04-21 paper/paper.qmd
update Data section

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 This paper uses data obtained from the open database, WHO Mortality database, which exhibits information on mortality causes of deaths filtered by country, sex, year, and age group.
-The data used in this paper were the Mortality causes in Switzerland in the years 2010,2015, and 2019 as it gives a general idea of the mortality causes from 2010 to 2019, filtered by age group.
+The data used in this paper were the Mortality causes in Switzerland in the years 2010, 2015, and 2019 as it gives a general idea of the mortality causes from 2010 to 2019, filtered by age group.
 
@@
 he data utilized for this paper is downloaded manually through the WHO Mortality Database.
-All the data analysis was done through R programming language (@citeR) with the aid of the following packages:rstanarm (@rstanarm), tidyverse (@tidyverse), knitr (@knitr), modelsummary (@modelsummary), ggplot2 (@ggplot2), dplyr (@dplyr), parameters (@Parameters), broom (@broom), broom.mixed (@broommixed).
+All the data analysis was done through R programming language (@citeR) with the aid of the following packages:rstanarm (@rstanarm), tidyverse (@tidyverse), knitr (@knitr), modelsummary (@modelsummary), ggplot2 (@ggplot2), dplyr (@dplyr), parameters (@parameters), broom (@broom), broom.mixed (@broommixed).
 
@@
 ```
-```
 

### 2024-04-21 paper/paper.qmd
Update typos and references

@@
 title: "Top Mortality causes in young adults and middle age adults in Switzerland"
-subtitle: "Analyzing Switzerland Mortality causes data in 2010, 2015, adn 2019 from WHO database"
+subtitle: "Analyzing Switzerland Mortality causes data in 2010, 2015, and 2019 from WHO database"
 author: 
@@
 ## Clean Data
-he data utilized for this paper is downloaded manually through the WHO Mortality Database.
+The data utilized for this paper is downloaded manually through the WHO Mortality Database.
 All the data analysis was done through R programming language (@citeR) with the aid of the following packages:rstanarm (@rstanarm), tidyverse (@tidyverse), knitr (@knitr), modelsummary (@modelsummary), ggplot2 (@ggplot2), dplyr (@dplyr), parameters (@parameters), broom (@broom), broom.mixed (@broommixed).
@@
 Notably, the leading cause of death for the whole population in Switzerland is Ischaemic heart disease with approximately 11,000 cases per year since 2010.
-(“Coronary Heart Disease in Switzerland” 2020).
+(@worldlife2020coronary).
 This disease is not only common in Switzerland, but it is also common worldwide, It results from reduced blood flow to the heart muscle due to narrowed or blocked coronary arteries.
 Some risk factors for this disease include high blood pressure, high cholesterol, smoking, diabetes, and obesity.
-It is reported that IHD is responsible for approximately 16% of total deaths worldwide1.(World 2020).
+It is reported that IHD is responsible for approximately 16% of total deaths worldwide.(World 2020).
 
@@
 Additionally, approximately 21% of smokers consumed at least 20 cigarettes per day.
-Overall, 6% of the Swiss population aged 15 and over were smokers.
-(“Level of Use - Swiss Association for Tobacco Control” 2016).
-It is shown that smokers are 2 or 4 times more likely to develop IHD than non-smokers.
+Overall, 6% of the Swiss population aged 15 and over were smokers.(@swiss2016tobacco) It is shown that smokers are 2 or 4 times more likely to develop IHD than non-smokers.
 With these statistics in mind, policymakers passed regulations such as banning smoking indoors and banning selling tobacco to people under 18.
@@
 Additionally, half of the patients were aged 18 to 34, and 65% of them were reported to have a difficult socioeconomic situation.
-(SWI swissinfo.ch 2019).
+(@swissinfo2019selfharm).
 In 2016, Switzerland launched a national action plan on self- harm/ suicide prevention with some key measures: raise awareness about these issues, provide accessible help, recognize early behavior of self-harm and timely intervene, reduce means of suicide and self-harm, etc.
@@
 Additionally, we can also compare the mortality causes and trends in Switzerland with those in other countries with similar socioeconomic conditions.
-As a result, we can identify the unique factors that make Switzerland one of the countries with the longest life expectancy.
+As a result, we can identify the unique factors that make Switzerland one of the countries with the highest life expectancy.
 
